\mythesistitle

### 2011-11-19 src/synopsis.tex
a whole bunch of changes. - put up the peanut dfw story. added images etc needed. i think some reorg of files were done too - organized the bib file chronologically. and fixed all the TBDs (except bl62 which is elusive and probably i'll delete it.

give some basic gyan about the problem itself? or screw it move to
section directly

@@

A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
  \Sc\ and \Lu\ enroll at the
% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
% enroll at the
{\em Wallace Studies Institute} for a liberal arts programme.  As part
of their semester thesis, they pick a body of work to study and form
the namesake study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}. A student
will be in at least one study group and may be in more than one. For
instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
while \Wo\ studies only \BBB.

@@

An intuitive approach to this problem would be to first find the paths
that each study group decides to inhabit and then refine the
allocation to individual students. One possible allocation of
exclusive routes to study groups is illustrated in
Fig~\ref{fig:streetmap+path}.

\begin{figure}[h]
  % \centering
  (a) \includegraphics[scale=0.3]{../img/a1.pdf}\label{fig:streetmap}
  (b) \includegraphics[scale=0.3]{../img/a2.pdf}\label{fig:streetmap+path}
  \caption{(a) {\residenceblock} street map (b) Street map with study
    group routes allocated. Routes are color coded as follows: red for
    \textcolor{red}{\em \LLL} group, blue for \textcolor{blue}{\em
      \GGG} group, orange for \textcolor{YellowOrange}{\em \BBB} group,
    green for \textcolor{green}{\em \TTT} group}
\end{figure}

@@

%\chapter{\MakeUppercase{Introduction}}
\chapter{INTRODUCTION}
Consecutive ones property (COP) of binary matrices is a widely studied
combinatorial problem. The problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its $1$s occur
consecutively. If this is possible the matrix is said to have the COP.
\begin{figure}[h]
  \centering

  \caption{(i) matrix with COP (ii) without COP}
  \label{fig:cop-matrix}
\end{figure}

This problem has several practical applications in diverse fields
including scheduling \cite{hl06}, information retrieval \cite{k77} and
computational biology \cite{abh98}.  Further, it is a tool in graph
theory \cite{mcg04} for interval graph recognition, characterization
of Hamiltonian graphs, and in integer linear programming
\cite{ht02,hl06}.

\tnote[a]{define ``COP permutation''}

\section{Background}
Different aspects of COP were studied as part of this thesis. They are
described in the following subsections.

\subsection{Consecutive Ones Testing}
The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  Since then several algorithms have been
invented; PQ trees \cite{bl76}, variations of PQ trees
\cite{mm09,wlh01,wlh02,mcc04}, ICPIA \cite{nsnrs09} are some of them.

\cite{bl76} showed for the first time that COP testing (COT) is indeed
polynomial time solvable by describing the construction of a data
structure called PQ trees. This paper draws on the close relationship
of COP matrices to interval graphs. A PQ tree of a matrix is one that
stores all row (column) permutations of the matrix that give the COP
orders (there could be multiple) of the matrix. This is constructed using an elaborate procedure in the linear
time algorithm described in \cite{bl76}.\tnote[a]{verify if it is
  linear}. Since this result, several improvements have been made to
the procedural complexity of the algorithm itself and also derivative
data structures that are easier to construct. PQR trees is one such
data structure \cite{mm96}, \cite{mpt97}\tnote{also, mpt97 $==$
  elaborate version of mm96}.  \cite{tm03} describes an improved
version of PQR trees. \tnote[a]{improvement in terms of
  what?}. \cite{h01a} describes the simpler algorithm for
COT. \tnote[a]{This result first apprearted in proc ISAAC 92} Hsu also
invented PC trees \cite{h01b} which is claimed to be much easier to
implement. \cite{nsnrs09} describes a characterization of consecutive
ones property solely based on the cardinality properties of the set representations
of the columns (rows); every column (row) is equivalent to a set that
has the row (column) indices of the rows (columns) that have one
entries in this column (row). This is interesting and relevant,
especially to this thesis because it simplifies COT to a great
degree. \tnote[a]{because it reduces the search space for the solution
  from set of all permutations of rows to set of all-- get this
  figured out!}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote[a] {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co97}.
\tnote[a]{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}

@@

% \noindent
To elaborate on what was mentioned earlier about \cite{nsnrs09}, the
problem of COT can be seen as a constraint satisfaction problem
involving a system of sets from a universe as follows. Every column of
the binary matrix can be converted into a set of integers which are
the indices of rows with $1$s in that column. It is apparent that if
the matrix has COP, then constructing such sets after applying the COP
permutation to the matrix will result in all sets with consecutive
integers. In other words, after application of COP reordering, the
sets are intervals. Indeed the problem now becomes finding interval
assignments to a given set system such that there exists a permutation
of the universe (set of row indices) which converts each set to its
assigned interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the rows.  They show that for each set, the cardinality
of the interval assigned to it must be same as the cardinality of the
set, and the intersection cardinality of any two sets must be same as
the intersection cardinality of the corresponding intervals.  While
this is obviously a necessary condition, this result shows this is
also sufficient.  \cite{nsnrs09} calls such an interval assignment an
Intersection Cardinality Preserving Interval Assignment (ICPIA).  This
paper generalizes the idea from \cite{wlh02} of decomposing a given
binary matrix into prime matrices for COT and describes an algorithm
to test if an ICPIA exists for a given set system.

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
\cite{cy91}\tnote[a]{is this parallel?}

@@

\subsection{Matrix modification to attain COP or ``almost'' COP}
\tnote[a]{THESIS: results and survey from \cite{d08phd} Chapters 2 and
  3}

So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP. Nevertheless, the COP is
not arbirary and there are several interesting problems when a matrix
does not have COP but is ``close'' to having COP or can be altered to
have COP.

@@

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete. \tnote[a]{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hgxx} \tnote[a]{get year XX} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem. \tnote[a]{THESIS: elaborate - what are
  the results?}.

@@

\tnote[a]{THESIS: where should this go? \cite{tz03} (approx submatrix with COP sparse matrices)}

@@

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{kou77}. \tnote[a]{THESIS: what is the reduction?}

\tnote[a]{THESIS: 
-- {\em other problems similar to COP}
  \begin{enumerate}
  \item \cite{ckl96} (ILP, circ ones, one drop)
  \item \cite{th99} (generalization of COP - minimax, biotonic column) Tucker
  \end{enumerate}
}

\subsection{Graph Isomorphism}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the
ordering\cite{gh64}. This clearly means that a graph $G$ is an
interval graph if and only if \tnote[a]{verify from paper the
  statement of claim.} maximal clique vertex incidence matrix of $G$
has COP.  Also, maximal cliques of any chordal graph can be enumerated
in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
results to give the first polynomial time algorithm for
COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
is convex (on $R$) if and only if its half adjacency matrix has COP on
rows.  The results in \cite{bl62} on COT are based on the result that
interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
  Tucker's?}

\subsection{Logspace complexity of graph canonization using COP}

Canonization is an important tool in graph isomorphism. Invention of a
deterministic method of canonization for any class of graphs naturally
results in an algorithm for isomorphism. All that is required is to
check if the canons of two graphs are the same. Thus complexity of
graph isomorphism can be studied by studying canonization
methods. While general graph ismorphism remains elusive in terms of
hardness, canonization has been studied for smaller classes of graphs
thus giving complexity/hardness results for them.

\cite{kklv10} proved that interval graph canonization is also in
logspace thus drawing logspace conslusions about COT. 
% This result is
% built on the fact that 
Interval graphs are FP+C (fixed point with
counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
this implies that it captures PTIME. \tnote[a]{elucidate} This result
along with that of undirected graph reachability being in logspace
\cite{r08}, \cite{kklv10} proved their logspace result.\\

\tnote[a]{THESIS: -- {\em \cite{dgnw06} (min membership hitting set + COP)}}

\section{Motivation, Objective and Scope}

A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote[a]{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

@@

%\noindent 
\section{Summary of Results}
Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.

@@

\section{Organization of document} 
\tnote[a]{THESIS: UNCOMMENT} 
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

%\chapter{\MakeUppercase{Results}}
\chapter{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

@@

%\section{Conclusions}

>>>>

\tnote{give some basic gyan about the problem itself? or screw it move to
section directly}

@@

\def \WSI {{\em Wallace Studies Institute}}

A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
  \Sc\ and \Lu\ enroll at the
% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
% enroll at the
{\WSI} for a liberal arts programme.  As part
of their semester thesis, they pick a body of work to study and form
the namesake study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}. A student
will be in at least one study group and may be in more than one. For
instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
while \Wo\ studies only \BBB.

@@

An intuitive approach to this problem would be to first find the paths
that each study group decides to inhabit and then refine the
allocation to individual students. A feasible allocation of exclusive
routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
and the students' allocation of apartments that obey's this route
allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
algorithmically computed is the focus of this thesis.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
  \caption{{\residenceblock} street map}
  \label{fig:streetmap}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
  \caption{{\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{\bf L} group, blue for \textcolor{blue}{\bf G}
    group, orange for \textcolor{YellowOrange}{\bf B} group, green for
    \textcolor{green}{\bf T} group}
  \label{fig:streetmappath}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
  \caption{Individual allocation of apartments to students in
    {\residenceblock} that meets the requirements stated before.}
  \label{fig:streetmappathpeople}
\end{figure}

If all the apartments are on the same street or if they are all lined
up on a single path, the street map becomes a tree that is just a
path. Then the problem becomes what is called an {\em interval
  assignment problem}. The idea of interval assignment may not be
obvious here; hence to see this, consider a different problem in
{\WSI} where the classes for these study groups courses need to be
scheduled during a day (or a week or any time period). Each study
group has a bunch of courses associated with it some of which may be
shared by two or more study groups. It is mandatory that a student who
is a member of a study group takes all the courses associated with
that group. There are slots during the day for classes to be held and
the problem is to allocate class slots to courses such that all the
classes of a study group are consecutive. It is debatable if this will
not hamper the attention span and memory retention rate of the students but
that is, regrettably, out of the scope of this thesis. The parallels between this
class allocation problem and the accommodation problem can be seen as
follows. The set $U$ here, are the courses offered (say Course 101 ``Post
modernism in Wallace's work'', Course 102 ``A study on fragmented prose
method'' and so on). The collection $\cF$ is the set of study groups but the
study groups are filled by course IDs and not students like in the
earlier example. For instance, Course 101 is mandatory for all study
groups {\bf L, G, B, T} and Course 102 is mandatory for only the {\bf B}
group) and so on. The sequence of class slots for the day (or week or
any time period) is analogous to the street map in the accommodation
problem. It is quite obvious now why this version of the problem
(where the ``target graph'' is a path and not any tree) is called an
interval assignment problem.

\tnote{TBD: BEGIN REVIEW FROM BELOW}

The interval assignment problem to a set system is equivalent to the
consecutive ones problem (COP) in binary matrices\cite{wlh02,
  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its $1$s occur
consecutively. If this is possible the matrix is said to have the COP.
Consecutive ones property (COP) of binary matrices is a well
researched combinatorial problem and has several positive results on
tests for it and computing the COP permutation (i.e. the class slot
allocation to course IDs in the above illustration) which will be
surveyed later in this document. Hence we are interested in extensions
of COP, more specifically, the extension of interval assignment
problem to tree path assignment problem (which is illustrated by the
study group accommodation problem).

\tnote[a]{THESIS: UNCOMMENT} \tnote{include the rest of the sections as
  the organization of the doc}
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
% Section~\ref{sec:feasible} documents the
% characterization of a feasible path labeling and finally,
% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
% find the tree path labeling of a given set system from a given
% $k$-subdivided tree.

Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
Chapter~\ref{chap:results} presents a summary of our results on the
tree path labeling problem.

@@

As descirbed earlier, a matrix with COP is one whose rows (columns)
can be rearranged so that the 1s in every column (row) are in
consecutive rows (columns).

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.4]{../img/0_cop_nocop.png}

  \caption{{\em Example for the COP:} The matrix on the left has the
    COP because by permuting its columns (labeled with $c_1$-$c_4$)
    one can obtain the matrix shown in the middle where the 1s in each
    row appear consecutively. The matrix on the right, in contrast,
    does not have the COP \cite{at72,d08phd}.}
  \label{fig:cop-matrix}
\end{figure}

COP in binary matrices has several practical applications in diverse
fields including scheduling \cite{hl06}, information retrieval
\cite{k77} and computational biology \cite{abh98}.  Further, it is a
tool in graph theory \cite{mcg04} for interval graph recognition,
characterization of Hamiltonian graphs, planarity testing \cite{bl76}
and in integer linear programming \cite{ht02,hl06}.

The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  Since then several algorithms have been
invented; PQ trees \cite{bl76} gave the first linear time
algorithm. Variations of PQ trees \cite{mm96,wlh01,wlh02,mcc04} and
ICPIA \cite{nsnrs09} further improve either the time complexity and/or
the complexity of implementation.

\cite{bl76} describes a linear time construction of a data structure
called PQ trees. This paper draws on the close relationship of COP
matrices to interval graphs. A PQ tree of a matrix is one that stores
all row (column) permutations of the matrix that give the COP orders
(there could be multiple) of the matrix. This is constructed using an
elaborate linear time procedure. Since this
result, several improvements have been made to the procedural
complexity of the algorithm itself and also derivative data structures
that are easier to construct. PQR trees is one such data structure
\cite{mm96}, \cite{mpt98}\tnote{also, mpt98 $==$ elaborate version of
  mm96}.  \cite{tm05} describes an improved version of PQR
trees. \tnote[a]{improvement in terms of what?}\cite{wlh02} describes
the simpler algorithm for COT. \tnote[a]{This result first apprearted
  in proc ISAAC 92} Hsu also invented PC trees \cite{wlh01} which is
claimed to be much easier to implement. \cite{nsnrs09} describes a
characterization of consecutive ones property solely based on the
cardinality properties of the set representations of the columns
(rows); every column (row) is equivalent to a set that has the row
(column) indices of the rows (columns) that have one entries in this
column (row). This is interesting and relevant, especially to this
thesis because it simplifies COT to a great degree. \tnote[a]{because
  it reduces the search space for the solution from set of all
  permutations of rows to set of all-- get this figured out!}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote[a] {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co98}.
\tnote[a]{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}

@@

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
\cite{ly91}\tnote[a]{is this parallel?}

@@

{\bf Matrices with no COP}: So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP. Nevertheless, the COP is
not arbirary and there are several interesting problems when a matrix
does not have COP but is ``close'' to having COP or can be altered to
have COP.

@@

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete. \tnote[a]{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hg02} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem. \tnote[a]{THESIS: elaborate - what are
  the results?}.

@@

\tnote[a]{THESIS: where should this go? \cite{tz04} (approx submatrix with COP sparse matrices)}

@@

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{k77}. \tnote[a]{THESIS: what is the reduction?}

\tnote[a]{THESIS: 
-- {\em other problems similar to COP}
  \begin{enumerate}
  \item \cite{ckl96} (ILP, circ ones, one drop)
  \item \cite{th98} (generalization of COP - minimax, biotonic column) Tucker
  \end{enumerate}
}

{\bf Motivation: } A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote[a]{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

@@

How the problem of COT is related to the constraint statisfaction
problem of interval assignment to sets in a system can be seen as
follows. Every column of the binary matrix can be converted into a set
of integers which are the indices of rows with $1$s in that column. It
is apparent that if the matrix has COP, then constructing such sets
after applying the COP permutation to the matrix will result in all
sets with consecutive integers. In other words, after application of
COP reordering, the sets are intervals. Indeed the problem now becomes
finding interval assignments to a given set system such that there
exists a permutation of the universe (set of row indices) which
converts each set to its assigned interval. The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the intersection cardinality of the
corresponding intervals.  While this is obviously a necessary
condition, this result shows this is also sufficient.  \cite{nsnrs09}
calls such an interval assignment an Intersection Cardinality
Preserving Interval Assignment (ICPIA).  This paper generalizes the
idea from \cite{wlh02} of decomposing a given binary matrix into prime
matrices for COT and describes an algorithm to test if an ICPIA exists
for a given set system.

Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.

@@

\section{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

@@

\chapter{Conclusion}
\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[a]{THESIS: peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the
ordering\cite{gh64}. This clearly means that a graph $G$ is an
interval graph if and only if \tnote[a]{verify from paper the
  statement of claim.} maximal clique vertex incidence matrix of $G$
has COP.  Also, maximal cliques of any chordal graph can be enumerated
in polytime $O(m+n)$ \tnote[a]{citation?!!}.  \cite{fg65} uses these
results to give the first polynomial time algorithm for
COT. \tnote[a]{THESIS: check. how do they use it?}  A bipartite graph
is convex (on $R$) if and only if its half adjacency matrix has COP on
rows.  The results in \cite{bl76}
%\cite{bl62} 
on COT are based on the result that
interval graphs are AT-free chordal graphs. \tnote[a]{the latter being
  Tucker's?}

{\bf Complexity challenges:} Canonization is an important tool in graph isomorphism. Invention of a
deterministic method of canonization for any class of graphs naturally
results in an algorithm for isomorphism. All that is required is to
check if the canons of two graphs are the same. Thus complexity of
graph isomorphism can be studied by studying canonization
methods. While general graph ismorphism remains elusive in terms of
hardness, canonization has been studied for smaller classes of graphs
thus giving complexity/hardness results for them.

\cite{kklv10} proved that interval graph canonization is also in
logspace thus drawing logspace conslusions about COT. 
% This result is
% built on the fact that 
Interval graphs are FP+C (fixed point with
counting) definable \tnote[a]{elucidate} and \cite{l10} showed that
this implies that it captures PTIME. \tnote[a]{elucidate} This result
along with that of undirected graph reachability being in logspace
\cite{rei08}, \cite{kklv10} proved their logspace result.\\
